---
title: "Stat341_Assignment1"
author: "(20726577)"
date: "24/01/2021"
output: pdf_document
---

## Question 4: Game of throne analysis

```{r}
# Read in a .csv file
df_got <- read.csv(file = "GOT.csv", header = TRUE)
head(df_got, n = 5)

```


```{r}
# order IMDb_Rating from highest to lowest
df_got[order(df_got$IMDb_Rating,decreasing = TRUE),]

```
### a

From the data we know, the directors directed the five most highly rated episodes are:

1st: David Nutter
2nd-4th: Miguel Sapochnik
5th: Matt Shakman

### b


```{r}
df_got[order(df_got$IMDb_Rating),]
```

From the data we know the name of the highest and lowest rated episodes are:

Highest: The Rains of Castamere
Lowest: The Iron Throne

### c

```{r}
# we pair director and write column and order the data by IMDb_Rating
aggregate(df_got[, 11], list(df_got$Director,df_got$Writer), mean)
```

From the data we know the name of the highest and lowest writer director pairing are:
Highest: Alex Graves & Bryan Cogman
Lowest: David Petrarca & Vanessa Taylor

### d

```{r}
# inital empty matrix
m = matrix(, nrow = 8, ncol = 10)

# loop through the data and add to matrix
for (i in 1:nrow(df_got)){
   m[df_got[i,]$ï..Season,df_got[i,]$Number_in_Season] = df_got[i,]$IMDb_Rating
}
print(m)
```

### e

```{r}
# I have not solve question 3 when doing this, I use build-in function for partial marks. Blue stands for lower, pink stands for higher.

m_heatmap <- heatmap(m, Rowv=NA, Colv=NA, col = cm.colors(256), scale="column")
```
From the diagram we can know season 8 have the worst rating in general.

### f


```{r}
# plot from csv
plot(df_got$IMDb_Rating, df_got$Episode_Number, main="IMDb_Rating vs.Episode_Number ",
   xlab="IMDb_Rating", ylab="Episode_Number", pch=16)

# plot every season average
tempArr = aggregate(df_got[, 11], list(df_got$ï..Season), mean)
xLast = 0
yLast = 0
for (i in 1:nrow(tempArr)){
   text(x = (tempArr[i,]$x), y = i*9,  labels = "X" , col = "red")
   if (xLast == 0){
      # if we are plotting the first average, we do not draw line segments
   }else{
      # connect with red line segments
      segments(xLast, yLast, (tempArr[i,]$x), i*9,
         col = "red")
   }
   xLast = (tempArr[i,]$x)
   yLast = i*9
}

# add legend
legend(4, 10, legend=c("each season average"),
       col=c("red"), lty=1:2, cex=0.8)

```

From this graph we can see the whole show is every popular and have a good reputation except the final season.


### g

We use sc function from question 2.

```{r}
# input rating data
popRating <- df_got$IMDb_Rating
y <- df_got$Episode_Number

# set sc with absolute function from class
sc_pos = function(y.pop, y, attr, ...) {
   N <-length(y.pop)+1 
   sapply( y,function(y.new) { abs(N*(attr(c(y.new, y.pop),...)- attr(y.pop,...)))  } )
}

par(mfrow=c(1,1))

# plot the data
plot(y,sc_pos(popRating, y, mean), type="l", lwd = 2,
     main="Sensitivity curve for the IMDb_Rating",xlab="Episode_Number",ylab="sensitivity")

abline(h=0, v=0, col="grey")
```
The largest influence episode is the last episode. It is because it has the lowest mark.

### h

We try 9 different alpha from 0 to 9.
```{r}
# input powerfun from class
powerfun <-function(x, alpha) {
   if(sum(x<=0)>0)stop("x must be positive")
   if(alpha==0)log(x)
   else if(alpha>0) {
      x^alpha
   }else -x^alpha
}

par(mfrow=c(3,3))
a =seq(0, 9, length.out=9)

for( i in 1:9 ) {
   hist(powerfun(df_got$IMDb_Rating+1, a[i]), col=adjustcolor("grey", alpha = 0.5),
        main=bquote(alpha==.(a[i])), xlab="x",  ylab="IMDb_Rating",breaks=50)
}

```
we have the best value of a as 6.75.

```{r}
par(mfrow=c(1,2))
a =seq(0, 6.75, length.out=2)

hist(df_got$IMDb_Rating, col="grey",
        main=bquote("original"), xlab="x",  ylab="IMDb_Rating",breaks=50)

hist(powerfun(df_got$IMDb_Rating+1, a[2]), col=adjustcolor("grey", alpha = 0.5),
        main=bquote(alpha==.(a[2])), xlab="x",  ylab="IMDb_Rating",breaks=50)
```


### i

```{r}
plot( df_got$IMDb_Rating, df_got$US_Viewers,
      pch = 16, cex=0.5,col=adjustcolor("black", alpha = 0.3),
      xlab = "x = IMDb_Rating", ylab = "y = US_Viewers",
      main = "Scatterplot IMDb_Rating vs. US_Viewers")
```

Since this is up on X and up on Y
Consider both a and b larger than 0.

We try 9 different alpha and beta from 0 to 4.
```{r}
par(mfrow=c(3,3), mar=2.5*c(1,1,1,0.1))

a =rep(c(0,2,4),each=3)
b =rep(c(0,2,4),times=3)
subdata = df_got[,c('IMDb_Rating','US_Viewers')]
subdata  =na.omit(subdata )

for(i in 1:9) {
   plot(powerfun(df_got$IMDb_Rating+1, a[i]),
        powerfun(df_got$US_Viewers+1, b[i]),
        pch = 16, cex=0.5,col=adjustcolor("black", alpha = 0.3),
        xlab = "", ylab = "",
        main = bquote(alpha[x]==.(a[i])~","~alpha[y]==.(b[i])))
}
```

a = 2, b = 2 looks more linear, so we adjust the function more.

```{r}

par(mfrow=c(3,3), mar=2.5*c(1,1,1,0.1))
a =rep(c(1.5,2,2.5),each=3)
b =rep(c(1.5,2,2.5),times=3)
subdata = df_got[,c('IMDb_Rating','US_Viewers')]
subdata  =na.omit(subdata )

for(i in 1:9) {
   plot(powerfun(df_got$IMDb_Rating+1, a[i]),
        powerfun(df_got$US_Viewers+1, b[i]),
        pch = 16, cex=0.5,col=adjustcolor("black", alpha = 0.3),
        xlab = "IMDb_Rating", ylab = "US_Viewers",
        main = bquote(alpha[x]==.(a[i])~","~alpha[y]==.(b[i])))
}
```



```{r}
par(mfrow=c(1,2))
a =seq(0, 6.75, length.out=2)

plot( df_got$IMDb_Rating, df_got$US_Viewers,
      pch = 16, cex=0.5,col=adjustcolor("black", alpha = 0.3),
      xlab = "x = IMDb_Rating", ylab = "y = US_Viewers",
      main = "Scatterplot IMDb_Rating vs. US_Viewers")


a =rep(c(2),each=1)
b =rep(c(2.5),times=1)
subdata = df_got[,c('IMDb_Rating','US_Viewers')]
subdata  =na.omit(subdata )

plot(powerfun(df_got$IMDb_Rating+1, a[1]),
        powerfun(df_got$US_Viewers+1, b[1]),
        pch = 16, cex=0.5,col=adjustcolor("black", alpha = 0.3),
        xlab = "x = IMDb_Rating", ylab = "y = US_Viewers",
        main = bquote(alpha[x]==.(a[1])~","~alpha[y]==.(b[1])))
```